SHIV: Reducing Supervisor Burden using Support Vectors to Efficiently Learn Robot Control Policies from Demonstrations [v21]

### 2015-09-14 paper.tex
Incoporated Anca's feedback

Rather than explicitly learning the cost function, like in Inverse Reinforcement Learning\adnote{cite}, and the dynamics model, and then using optimization to produce a policy for the task, in model-free LfD the robot learns the policy directly from supervisor examples mapping states to controls~\cite{argall2009survey}.

@@

\includegraphics[width=0.5\textwidth]{figures/teaser.pdf}
\caption{ 
SHIV  reduces the number of queries to the supervisor in our three test domains: a) a driving simulator where the goal is to learn a controller on HOG features extracted from synthetic images to keep a car on a polygonal track; b) push-grasping in clutter in a physics simulator, where the goal is to learn a controller from human demonstrations to grasp the green object while not letting the red square objects fall off the gray boundary representing a table; c) learning a controller for the Da Vinci Research Kit from a master surgeon's demonstrations of correcting a needle before suturing in surgery; d) normalized performance of SHIV and DAgger's policy on the grasping in clutter example versus states labeled \adbote{by the supervisor}.
}
\vspace*{-10pt}
\label{fig:teaser}
\end{figure}

LfD algorithms can be categorized as offline, where the robot only observes demonstrations, or online, where the robot interacts with the world and receives feedback from the supervisor.
In offline LfD, the robot learns the policy based on a batch of examples, and then executes it to achieve the task.  During execution, a small error can accumulate, leading the robot away from the region of the state space where it was given examples, leading to unrecoverable failures.  For example, a robot driving may be trained on examples driving safely down the center of a lane, but even slight deviations will \adnote{eventually} put the robot into states near the side of the road where its policy could  fail~\cite{pomerleau1989alvinn}. Ross and Bagnell showed the number of errors made by the robot, in the worst case,  can scale quadratically with the time horizon of the task~\cite{ross2010efficient}.

Online LfD addresses this issue by iteratively gathering more examples from the supervisor in states the robot encounters~\cite{grollman2007dogged,ross2010efficient,ross2010reduction}. One such algorithm, DAgger, learns a series of policies. At each iteration, the robot trains a policy based on the existing examples, then rolls out (executes) that policy, and the supervisor provides demonstrations for all states the robot visits. The new state/control examples are aggregated with the old examples for the next iteration. DAgger and related algorithms have been applied in a wide range of applications, from quadrotor flight to natural language to Atari games~\cite{NIPS2014_5421,duvallet2013imitation,ross2013learning}. In DAgger, \adnote{delted it was proved, unnecessary} under certain conditions, the number of errors scales only linearly with the time horizon of the task\adnote{cite}.

\adnote{there needs to be a lead up to this, like yay dagger is cool. however, }DAgger significantly increases the burden on the supervisor, who now labels all states that the robot visits during training.  Our goal is to reduce supervisor burden without degrading robot performance.

@@

\noindent\textbf{Active Learning from Demonstration}
Online LfD\adnote{been replacing with "LfD" for simplicity, no need to say the same expression 100 times in the paper especially since everyone calls it that} can be formulated as a stream-based active learning problem ~\cite{atlas1990training,cohn1994improving}\adnote{NO!! Your approach of asking only when risky instantiates stream based active learning. In general, online lfd is NOT an active learning problem, that is the proble we are fixing here! sorry for the "!!" but this is crucial}. In stream based active learning, the decision of whether to query the supervisor (for a label or control signal) or not is not over the entire state space (like in traditional pool-based active learning), but on states drawn one at a time from some data stream. In  online LfD, the\adnote{this} data stream is the states encountered during learning\adnote{encountered when the robot is executing the current best policy.}

Stream-based active learning has two ingredients: a stream, given here by executing the current best policy,\adnote{Aside from the data strea, stream-based active learning has another ingredient: a query selection method.} and a query selection method deciding whether or not to query. Typical query selection methods are estimator-centric, evaluating risk using the estimator, e.g. distance from the classification hyperplane \cite{tong2002support}, as in Fig. \ref{fig:support_example}(a); or query by committee  \cite{breiman1996bagging}, which uses a committee of hypothesized estimators that are trained on different subsets of the training data. Risk in query by committee is based on the level of agreement or disagreement among these hypotheses, with higher levels of disagreement for a state leading to higher chances of querying that state (Fig. \ref{fig:support_example}(b)).

@@

Our approach is based on the One Class SVM proposed by Scholk{\"o}pf et al., which estimates a particular quantile level set for the training data by solving a convex quadratic program  to the find support
vectors~\cite{scholkopf2001estimating}. The method has been theoretically shown to approximate the quantile levelset of a density estimate
asymptotically for correctly chosen bandwidth settings and in the case of a normalized Gaussian kernel function \cite{vert2006consistency}. 
In \cite{liu2014unsupervised}, the One Class SVM has furthermore been used for novelty detection in high-dimensional image data. To our knowledge the use of One Class SVM to online learning from demonstration has never been done before. \mlnote{Ken, wanted to our knowledge in the last part} \adnote{you can say this to me, but not really write it in a paper; I think that perhaps you want to say, and in a new paragraph so that it stands out, "Our contribution is the addition of stream-based active learning to online learning from demonstration. To do so, we use a query selection method based on the One Class SVM.}

\adnote{OK fiiiiiine. Then maybe To our knowledge, incorporating a query selection method for nonstationary distributions, such as the One Class SVM, into online LfD is a novel contribution.}

@@

\noindent\textbf{Modeling Choices and Assumptions}  We model the system dynamics as Markovian, stochastic, and stationary. Stationary dynamics meaning\adnote{this sentence is missing a predicate; Stationary dynamics occur when, given..}  given a state and a control, the probability of the next state doesn't\adnote{does not} change over time. Note this is different then\adnote{from} the non-stationary distribution over the states the robot encounters during learning. \adnote{very good addition here!}

We model the initial state as sampled from a distribution over the state space.
We assume a known state space and set of controls. We also assume access to a robot or simulator, such that we  can sample from the state sequences induced by a sequence of controls.   Lastly, we assume access to a supervisor who can, given a state, provide a control signal label. We assume the supervisor can be noisy and imperfect, noting that the robot cannot surpass the level of performance of the supervisor. \adnote{not in general just in our work}

\adnote{again, this would be better at the same time with introducing the basic notation, like states and controls, so that you can state mathematically what these mean; for instance, you restate the markovian modelling anyway}

\noindent\textbf{Policies and State Densities.}
Following conventions from control theory, we denote by $\mathcal{X}$ the set consisting of observable states for a robot task, consisting, for example, of \mlnote{We should discuss if we are going to consider images states or not}
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of allowable control inputs for the robot, which can be discrete or
continuous. We model dynamics as Markovian, such that the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$: 
$$p(\bx_{t+1}|\bu_{t},\bx_{t}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{t+1}|\bu_{t}, \bx_t)$$
We assume a probability density over initial states $p(\bx_0)$.
%We denote the probability density over the initial state also by $p:\mathcal{X}\to \mathbb{R}$.

@@

\noindent\textbf{Objective.} The objective of  policy learning is to find a policy that minimizes some known cost function $C(\hat{\tau}) = \sum^T_{t=1} c(\bx_t,\bu_t)$ of a trajectory $\hat{\tau}$. The cost $c:\mathcal{X}\times \mathcal{U}\to \mathbb{R}$ is typically user defined and task specific. 
For example, in the task of inserting a peg into a hole, a function on the distance between the peg's current and desired final state can
be considered \cite{levine2015end}.  \adnote{can be considered? like you'll allow it? maybe For example, in the task... , the cost function could be ...}

In our problem, we do not have access to the cost function itself. Instead, we only have access to 
a supervisor that can achieve a desire\adnote{d} level of performance on the task. \adnote{furthermore doesnt make sense here because the supervisor is providing these; say instead The supervisor provides the robot an initial set..} We are furthermore given
an initial set of $N$ stochastic demonstration trajectories $\lbrace \tilde{\tau}^1,...,\tilde{\tau}^N \rbrace$. 
which are the result of the supervisor applying this policy. This induces a training data set $\mathcal{D}$ of all state-control input pairs from the demonstrated trajectories.

We define a `surrogate' loss function \adnote{as in \cite{ross2010reduction}, }$l:\mathcal{U}\times \mathcal{U}\to \mathbb{R}$, which provides a distance
measure between any pair of control values. In the continuous case, we consider $l(\bu_0,\bu_1) = ||\bu_0-\bu_1||^2$,
while in the discrete case $l(\bu_0,\bu_1) = 1$ if $\bu_0 \neq \bu_1$ and $l(\bu_0, \bu_1)=0$ otherwise.

@@

 \vspace{-2ex}
\begin{align}\label{eq:LFD_obj}
\underset{\theta}{\min} \: E_{p(\bx|\theta)} [l(\pi_\theta(\bx),\tilde{\pi}(\bx))]
\end{align}
with $p(\bx|\theta)$ from Eq. \ref{eq:density}.\adnote{fix}

@@

\section{Risky and Safe States}
\adnote{I don't like the flow of this coming before the algorithm. You can understand the algorithm without understanding the specifics of risk, but here I read this and still don't know how the algorithm is going to work and it's like keeping me in (bad) suspense until the next section. It also breaks flow because the last section set up the optimization problem that shiv and dagger are solving}
\begin{figure*}[t]
\centering

@@

Providing correct control inputs for (or ``labeling'')  all states encountered at each
iteration can impose a large burden on the supervisor.\adnote{but you haven't described the algorithm yet so how am i supposed to know what that means; this is no longer intro land so stuff should be precise not handwavy; that is why this section doesnt fit here}
Instead of asking the supervisor for labels at all visited states, SHIV uses a measure of risk to actively decide whether a label is necessary.

@@

We then  compare our query selection method with those typically used in active learning. We show that for a non-stationary state distribution like ours, the notion of risk based on novelty and misclassified regions performs better than a confidence and query-by-committee based methods.\adnote{also: We also test whether SHIV actually benefits from measuring both forms of risk (novelty and ???).} We continue with a sensitivity analysis, which suggests that the performance of SHIV is robust to the choice of how risky the robot is allowed to be (the $\sigma$ parameter from Eq. \ref{eq:decision_func}).

>>>>

Rather than explicitly learning the cost function, like in Inverse Reinforcement Learning \cite{ng2000algorithms}, and the dynamics model, and then using optimization to produce a policy for the task, in model-free LfD the robot learns the policy directly from supervisor examples mapping states to controls~\cite{argall2009survey}.

@@

\includegraphics[width=0.5\textwidth]{figures/teaser.pdf}
\caption{ 
SHIV  reduces the number of queries to the supervisor in our three test domains: a) a driving simulator where the goal is to learn a controller on HOG features extracted from synthetic images to keep a car on a polygonal track; b) push-grasping in clutter in a physics simulator, where the goal is to learn a controller from human demonstrations to grasp the green object while not letting the red square objects fall off the gray boundary representing a table; c) learning a controller for the Da Vinci Research Kit from a master surgeon's demonstrations of correcting a needle before suturing in surgery; d) normalized performance of SHIV and DAgger's policy on the grasping in clutter example versus states labeled by the supervisor.
}
\vspace*{-10pt}
\label{fig:teaser}
\end{figure}

LfD algorithms can be categorized as offline, where the robot only observes demonstrations, or online, where the robot interacts with the world and receives feedback from the supervisor.
In offline LfD, the robot learns the policy based on a batch of examples, and then executes it to achieve the task.  During execution, a small error can accumulate, leading the robot away from the region of the state space where it was given examples, leading to unrecoverable failures.  For example, a robot driving may be trained on examples driving safely down the center of a lane, but even slight deviations will eventually put the robot into states near the side of the road where its policy could  fail~\cite{pomerleau1989alvinn}. Ross and Bagnell showed the number of errors made by the robot, in the worst case,  can scale quadratically with the time horizon of the task~\cite{ross2010efficient}.

Online LfD addresses this issue by iteratively gathering more examples from the supervisor in states the robot encounters~\cite{grollman2007dogged,ross2010efficient,ross2010reduction}. One such algorithm, DAgger, learns a series of policies. At each iteration, the robot trains a policy based on the existing examples, then rolls out (executes) that policy, and the supervisor provides demonstrations for all states the robot visits. The new state/control examples are aggregated with the old examples for the next iteration. DAgger and related algorithms have been applied in a wide range of applications, from quadrotor flight to natural language to Atari games~\cite{NIPS2014_5421,duvallet2013imitation,ross2013learning}. In DAgger, under certain conditions, the number of errors scales only linearly with the time horizon of the task\cite{ross2010reduction}.

However one drawback is that DAgger significantly increases the burden on the supervisor, who now labels all states that the robot visits during training.  Our goal is to reduce supervisor burden without degrading robot performance.

@@

\noindent\textbf{Active Learning from Demonstration}
We formulate applying active learning to Online Lfd as a stream-based active learning problem ~\cite{atlas1990training,cohn1994improving}. In stream based active learning, the decision of whether to query the supervisor (for a label or control signal) or not is not over the entire state space (like in traditional pool-based active learning), but on states drawn one at a time from some data stream. In  online LfD, this data stream is the states encountered when the robot is executing the current best policy.

Aside from the data stream, stream-based active learning has another ingredient: a query selection method. A query selection method is deciding whether or not to query. Typical query selection methods are estimator-centric, evaluating risk using the estimator, e.g. distance from the classification hyperplane \cite{tong2002support}, as in Fig. \ref{fig:support_example}(a); or query by committee  \cite{breiman1996bagging}, which uses a committee of hypothesized estimators that are trained on different subsets of the training data. Risk in query by committee is based on the level of agreement or disagreement among these hypotheses, with higher levels of disagreement for a state leading to higher chances of querying that state (Fig. \ref{fig:support_example}(b)).

@@

Our approach is based on the One Class SVM proposed by Scholk{\"o}pf et al., which estimates a particular quantile level set for the training data by solving a convex quadratic program  to the find support
vectors~\cite{scholkopf2001estimating}. The method has been theoretically shown to approximate the quantile levelset of a density estimate
asymptotically for correctly chosen bandwidth settings and in the case of a normalized Gaussian kernel function \cite{vert2006consistency}. 
In \cite{liu2014unsupervised}, the One Class SVM has furthermore been used for novelty detection in high-dimensional image data. To our knowledge, incorporating a query selection method for nonstationary distributions, such as the One Class SVM, into online LfD is a novel contribution.

@@

\noindent\textbf{Modeling Choices and Assumptions}  We model the system dynamics as Markovian, stochastic, and stationary. Stationary dynamics occur when, given a state and a control, the probability of the next state does not change over time. Note this is different from the non-stationary distribution over the states the robot encounters during learning.

We model the initial state as sampled from a distribution over the state space.
We assume a known state space and set of controls. We also assume access to a robot or simulator, such that we  can sample from the state sequences induced by a sequence of controls.   Lastly, we assume access to a supervisor who can, given a state, provide a control signal label. We additionally assume the supervisor can be noisy and imperfect, noting that the robot cannot surpass the level of performance of the supervisor.

\noindent\textbf{Policies and State Densities.}
Following conventions from control theory, we denote by $\mathcal{X}$ the set consisting of observable states for a robot task, consisting, for example, of 
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of allowable control inputs for the robot, which can be discrete or
continuous. We model dynamics as Markovian, such that the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$: 
$$p(\bx_{t+1}|\bu_{t},\bx_{t}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{t+1}|\bu_{t}, \bx_t)$$
We assume a probability density over initial states $p(\bx_0)$.
%We denote the probability density over the initial state also by $p:\mathcal{X}\to \mathbb{R}$.

@@

\noindent\textbf{Objective.} The objective of  policy learning is to find a policy that minimizes some known cost function $C(\hat{\tau}) = \sum^T_{t=1} c(\bx_t,\bu_t)$ of a trajectory $\hat{\tau}$. The cost $c:\mathcal{X}\times \mathcal{U}\to \mathbb{R}$ is typically user defined and task specific. 
For example, in the task of inserting a peg into a hole, a function on the distance between the peg's current and desired final state can
be used \cite{levine2015end}.

In our problem, we do not have access to the cost function itself. Instead, we only have access to 
a supervisor that can achieve a desired} level of performance on the task. The supervisor provides the robot an initial set
an initial set of $N$ stochastic demonstration trajectories $\lbrace \tilde{\tau}^1,...,\tilde{\tau}^N \rbrace$. 
which are the result of the supervisor applying this policy. This induces a training data set $\mathcal{D}$ of all state-control input pairs from the demonstrated trajectories.

We define a `surrogate' loss function as in \cite{ross2010reduction}, $l:\mathcal{U}\times \mathcal{U}\to \mathbb{R}$, which provides a distance
measure between any pair of control values. In the continuous case, we consider $l(\bu_0,\bu_1) = ||\bu_0-\bu_1||^2$,
while in the discrete case $l(\bu_0,\bu_1) = 1$ if $\bu_0 \neq \bu_1$ and $l(\bu_0, \bu_1)=0$ otherwise.

@@

 \vspace{-2ex}
\begin{align}\label{eq:LFD_obj}
\underset{\theta}{\min} \: E_{p(\bx|\theta)} [l(\pi_\theta(\bx),\tilde{\pi}(\bx))]
\end{align}
with $p(\bx|\theta)$ from Eq. \ref{eq:density}.

@@

\section{Risky and Safe States}

\begin{figure*}[t]
\centering

@@

Providing correct control inputs for (or ``labeling'')  all states encountered at each
iteration can impose a large burden on the supervisor.
Instead of asking the supervisor for labels at all visited states, SHIV uses a measure of risk to actively decide whether a label is necessary.

@@

We then  compare our query selection method with those typically used in active learning. We show that for a non-stationary state distribution like ours, the notion of risk based on novelty and misclassified regions performs better than confidence, query-by-committee based methods and the One Class SVM with out the hole carving out modification. We continue with a sensitivity analysis, which suggests that the performance of SHIV is robust to the choice of how risky the robot is allowed to be (the $\sigma$ parameter from Eq. \ref{eq:decision_func}).
